---
title: "biostatistics_assignment03"
author: "Kevin Jou"
date: "2023-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2)
library(ggplot2)
```

```{r}
wh <- read.csv("weight-height.csv",header=TRUE)
```

## 1)  Plot the distribution of heights in the group.

```{r}
#list distribution of rounded heights
rounded_heights <- round(wh$"Height", 0)
hist(rounded_heights, main="Distribution of Heights", xlab="Height (rounded)", ylab="Frequency")
```
## Is the distribution Gaussian ?
Yes, the shape of the distribution appears to follow that of a normal distribution.
## 3) If the distribution is Gaussian, does it mean there are no sub groups ?  
Not necessarily, it could mean that sub groups within the grouped distribution may follow a normal distribution as well, with the largest overlap of the subgroups appearing where the mean of the grouped distribution lies. 

## 4) Plot the distribution of heights of men in black and women in red on the same graph  and for the whole group (in blue)

```{r}
# Filter data for males and females
males <- subset(wh, wh$Gender == "Male")
females <- subset(wh, wh$Gender == "Female")

# Round heights for males and females
rounded_heights_male <- round(males$"Height", 0)
rounded_heights_female <- round(females$"Height", 0)

# Create histograms for males, females, and whole group
hist(rounded_heights, main="Distribution of Heights", xlab="Height (rounded)", ylab="Frequency", col="lightblue")
hist(rounded_heights_male, col="black", add=TRUE)
hist(rounded_heights_female, col="red", add=TRUE)
```
## Are the “Male” and “Female” distributions gaussian ?
In this case they do appear to both be gaussian for height.
## 5) Repeat 1-4 with weight instead of height
```{r}
rounded_weights <- round(wh$"Weight", 0)
hist(rounded_heights, main="Distribution of Weights", xlab="Weight (rounded)", ylab="Frequency")

# Round weights for males and females
rounded_weights_male <- round(males$"Weight", 0)
rounded_weights_female <- round(females$"Weight", 0)

# Create histograms for males, females, and whole group
hist(rounded_weights, main="Distribution of Heights", xlab="Height (rounded)", ylab="Frequency", col="lightblue")
hist(rounded_weights_male, col="black", add=TRUE)
hist(rounded_weights_female, col="red", add=TRUE)
```
## Is the distribution Gaussian ?
No, the shape of the distribution appears to follow that of a bimodal distribution.

## Are the “Male” and “Female” distributions gaussian ?
In this case they do appear to both be gaussian for weight, but are significantly far apart as to create the bimodal distribution when combined.
## 6) plot height (x-axis) versus weight (y-axis) for the group

```{r}
#Plot distribution of height vs. weight, one dot represents one individual
plot(wh$"Height", wh$"Weight", main="Height vs. Weight", xlab="Height", ylab="Weight", pch=16, col="blue")
```

## 7) plot height versus weight, separately for men (in black) and women (red).

```{r}
plot(males$Weight, males$Height, pch=16, col="black", xlab="Weight", ylab="Height", main="Weight vs. Height (Males)",xlim=c(40, max(males$Weight)), ylim=c(50, max(males$Height)))
points(females$Weight, females$Height, pch=16, col="red")
legend("topleft", legend=c("Males", "Females"), col=c("black", "red"), pch=16)

```
# 8) Discuss the following
   a) is height a good proxy for weight ?
   Height may act as a decent proxy for weight, depending on the margin of error that you wish to account for as there does appear to be a positive correlation between height and weight.

   b) Is the rate of increase in weight with height higher for men or for women ?
   It appears to be about equal for both genders, maybe higher in men but both genders appear to show very similar correlations between height and weight.

   c) Based on these plots and values for standard deviations and means, for which group (men versus women) is the height a better proxy for weight ?
   The men since they exhibit lower overall variability for height and weight while showing a correlation for height and weight.  This points torward a stronger basis for assuming better fitting of the data to males.


   d_ Would binning the data on height and weight and using a 3-d plot give a better picture ?
   It would allow for a third dimension which may better represent the real factors which can contribute to weight and height as well.  Such as age/other demographic information.
   
# 9) Use the dataset called iris.
   a) Describe the data through two interesting plots.
   b) Use the reshape2 package to show multiple panels in one plot
```{r}
data(iris)
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  labs(title="Scatter Plot of Sepal Length vs Sepal Width", x="Sepal Length", y="Sepal Width")

ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point() +
  labs(title="Scatter Plot of Sepal Length vs Petal Length", x="Sepal Length", y="Petal Length")
```

```{r}
library(reshape2)
library(ggplot2)
melted_iris <- melt(iris, id.vars="Species")
ggplot(melted_iris, aes(x=variable, y=value, fill=Species)) +
  geom_dotplot(binaxis='y', stackdir='center', position=position_dodge(0.75)) +
  facet_wrap(~variable, scales="free_y") +
  labs(title="Panel Plot of Iris Measurements", x="Measurement", y="Value") +
  theme(legend.position="top")  # Move legend to the top
```

