---
title: "biostatistics_week12"
author: "Kevin Jou"
date: "2023-12-27"
output: html_document
---

```{r}
data(warpbreaks)

# Fitting a Poisson model
poisson_model <- glm(breaks ~ tension + wool, data = warpbreaks, family = poisson)

# Fitting a quasi-Poisson model
quasi_poisson_model <- glm(breaks ~ tension + wool, data = warpbreaks, family = quasipoisson)

# Comparing the models
summary(poisson_model)
summary(quasi_poisson_model)

```

```{r}
library(jtools)

# Create summaries for the models
poisson_summary <- summ(poisson_model)
quasi_poisson_summary <- summ(quasi_poisson_model)

# Plotting model variables for Poisson model
plot_summs(poisson_summary)

# Plotting model variables for quasi-Poisson model
plot_summs(quasi_poisson_summary)

```

Analysing the resulting graphs from the poisson and quasipoisson link function, the quasipoisson link function appears to have a higher region of error, which would mean that in this case the data would more closely fit a poisson distribution than a quasi-poisson distribution.  For other datasets quasi-poisson may be better depending on the resulting dispersion in your data.

