---
title: "stemID"
author: "Kevin Jou"
date: "2023-08-17"
output: html_document
---
# StemID
StemID analyses takes output from RaceID2 or from conversion of a seurat object.  In our case since we used scanpy, we must use the original count matrix from the ductal subset as our input.

## Load Required R Libraries
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(RaceID)
library(Seurat)
library(Matrix)
```
RaceID3 (J. S. Herman, Sagar, and Grun 2018): pre-processing, clustering and outlier identification
StemID2 (J. S. Herman, Sagar, and Grun 2018): lineage tree inference (dependend on a prior RaceID3 analysis)

## Create Scseq object for RaceID2 and StemID Analysis
```{r pressure, echo=FALSE}
table <- read.csv("C:/Users/Kevin/Documents/Summer semester/duct_count.csv", header=TRUE)
genes <- table[,1]
genefiltered <- genes[!duplicated(genes)]
filteredtable <- table[match(genefiltered, table[,1]), ]
rownames(filteredtable) <- genefiltered
filteredtable <- filteredtable[,-1]
test <- CreateSeuratObject(filteredtable, project="Ducts")
sc <- Seurat2SCseq(test)
sc <- compdist(sc,metric="pearson")
sc <- clustexp(sc)
plotsaturation(sc,disp=TRUE)
plotjaccard(sc)
```
Table format output from python has some issues, such as duplicate row entries due to how strings are converted.  R requires us to remove these features and assign row names before accepting the data frame as a matrix which can then be loaded as a seurat object to keep its gene and cell annotations before converting to a SCseq object using the Seurat2SCseq function in RaceID.

Generated graphs show default cluster number is an overestimate, leading to instability as indicated by low Jaccard’s similarity. Based on visual inspection of the average within-cluster dispersion as a function of the cluster number, we manually set the cluster number to 14 without recomputing the saturation behavior.

```{r}
sc <- clustexp(sc,cln=14,sat=FALSE)
```
```{r}
sc <- findoutliers(sc)
plotoutlierprobs(sc)
```
```{r}
sc <- compumap(sc)
sc <- compfr(sc)
sc <- comptsne(sc)
```
Dimensional reduction methods are perfomed on dataset for use in StemID tree analyses.

```{r}
ltr <- Ltree(sc)
ltr <- compentropy(ltr)
```

```{r}
ltr <- projcells(ltr,cthr=5,nmode=TRUE, knn=3)
ltr <- projback(ltr,pdishuf=30)
```
StemID links are generated using Fruchterman-Rheingold representation and then computation is performed 30(>20 ideally) times to ensure quality of data.

```{r}
ltr <- lineagegraph(ltr)
```

```{r}
ltr <- comppvalue(ltr, pthr=0.05)
plotgraph(ltr,scthr=0.2,showCells=FALSE,showMap=TRUE)
```

## Plot of potential stem cell clusters residing in the ductal cluster.  Node color represents 
```{r}
x <- compscore(ltr,scthr=0.2)
plotlinkscore(ltr)
projenrichment(ltr)
```
Most clusters identified in the ductal portion appear to be showing high scores for stem cell identity based on the prediction algorithm, however it is difficult to ascertain the significance of these results due to the high entropy signified by the color of the vertices.
```{r}
genes <- c("NKX6-1", "SOX9", "PDX1", "ITGA6", "PROM1")
fractDotPlot(sc, genes, cluster=c(2,3,4,6,7,8,9,10,11,13,15,16,17,18,20,21,22,23), zsc=TRUE)
```
The dotplot here shows that in this case, the cells of interest may actually be the "low" scoring clusters 2,12, and 16 as 2 and 12 both show a higher proportion of ITGA6, PROM1 expression compared to other clusters.  Cluster 19 had the most Sox9 expression but also scored lower on the stemID analysis.  This may point to a difference in the nature of the progenitor like cells compared to the known characteristics of stem cells.  Across most clusters there appear to be groups of enrichment, which are positive and may represent a higher proportion than usual.







